---
title: "main_v5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in data

```{r setup, include=FALSE}
#install.packages('moments')
library(readxl)
library(moments)
library(ggplot2)

#filepath_ellie = "/Users/EllieLiu/Desktop/Monthly_returns.xlsx"
filepath_haocheng = '/Users/haochengtang/Desktop/MIT/Fall2018/Proseminar/data/clean\ data/Monthly_returns.xlsx'
#filepath_viktor = '/Users/hermannviktor/Python/PycharmProjects/15.451-Proseminar-IM/Excel/Monthly_returns.xlsx'

dt1 <- read_excel(filepath_haocheng)

```

Function 1:
input: 1. hitorical return data
       2. SORR 
       3. number of paths needed

output: return paths for equity and equity
        path_equity and path_FI are matrices with 342 rows (days of period) and n columns (number of path)
        
```{r setup, include=FALSE}
dt = dt1[, c(2,3,4)]
dt = dt[sample(1:nrow(dt), size = 360, replace = FALSE),]
colnames(dt) <- c("equity","FI","Inflation")

base = dt[order(dt$equity),]
n = nrow(base) 
equity_return = base$equity
FI_return = base$FI
Inflation = base$Inflation

size_bad_return = floor(0.3*n)
size_ok_return = n-size_bad_return
Simulated_Path_Given_SORR =
  function(sorr,m){
    index_bad = 1:floor(0.3*n) #108 periods
    index_ok = (floor(0.3*n)+1):n
    rand_index = matrix(0, nrow = n, ncol = m)
    path_equity = matrix(0, nrow = n, ncol = m)
    path_FI = matrix(0, nrow = n, ncol = m)
    path_Inflation = matrix(0, nrow = n, ncol = m)
    
    for(i in 1:m){
      left_bad = sample(index_bad, size = floor(sorr*size_bad_return))
      right_bad = setdiff(index_bad, left_bad)
      left_ok = sample(index_ok, size = (1/2*n - floor(sorr*size_bad_return)))
      right_ok = setdiff(index_ok, left_ok)
      left = c(left_bad, left_ok)
      right = c(right_bad, right_ok)
      
      #rand_index[1:(1/2*size_bad_return),i] = left_bad
      #rand_index[(1/2*n+1):(1/2*n+sorr*size_bad_return),i] = right_bad
      #rand_index[(1/2*size_bad_return+1):(1/2*n),i] = left_ok
      #rand_index[(1/2*n+1/2*size_bad_return+1):n,i] = right_ok
      
      rand_index[(1:(1/2*n)),i] <- sample(left, size = length(left))
      rand_index[((1/2*n+1):n),i] <- sample(right, size = length(right))
    }
    
    for (i in 1:m) {
      for (j in 1:n) {
        path_equity[j,i] = equity_return[rand_index[j,i]]
        path_FI[j,i] = FI_return[rand_index[j,i]]
        path_Inflation[j,i] = Inflation[rand_index[j,i]]
      }
    }
    return(list(path_equity,path_FI,path_Inflation))
  }

gene_path <- Simulated_Path_Given_SORR(0.5,100)
path_equity <- gene_path[[1]]
path_FI <- gene_path[[2]]
path_Inflation <- gene_path[[3]]
```

Function 2:
generate glidepath
mode 1: constant weight
mode 2: constantly changing weight

```{r}
gene_glidepath = function(mode,constant,start,end) {
  # mode = 1; constant weight for equity 
  glidepath = matrix(0,nrow = n, ncol = 2)
  if (mode == 1) {
    glidepath[,1] = constant
    glidepath[,2] = 1-constant
  }
  
  # mode = 2; constantly changing weight 
  if (mode == 2) {
    rate = (end - start)/360  #changing rate
    glidepath[,1] = seq(from = start+rate, to = end, by = rate)
    glidepath[,2] = 1-glidepath[,1]
  }
  return(glidepath)
}
```


Function 3:
input: 1. given return paths of equity, fixed income, inflation for the period of time
       2. asset allocation glidepath
       3. retirement fund starting balance and withdraw rate
       4. number of paths

output: average coverage ratio for given glidepath and SORR

```{r setup, include=FALSE}
SORR_to_coverage = 
  function(path_equity,path_FI,path_Inflation,glidepath_one,e,x,n,m) {
  withdraw = matrix(0,nrow = n, ncol = m)
  
  init = e*x/12
  withdraw[1,] = init
  
  for (i in 2:nrow(dt)) {
    for (j in 1:ncol(path_Inflation)) {
      withdraw[i,j] = withdraw[i-1,j] * (1+path_Inflation[i,j])
    }
  }
  
  wealth = matrix(0,nrow = n, ncol = m)
  wealth[1,] = e
  
  #calculate wealth position at specific time point after withdraw
  for (j in 1:m) {
    for (i in 2:n) {
      wealth[i,j] = (wealth[i-1,j] * ((1+path_equity[i,j])*glidepath[i,1] + (1+path_FI[i,j])*glidepath[i,2])) - withdraw[i,j]   
    }
  }
  
  
  # calculate coverage ratio for every path
  # if wealth at 30 year is still positive, coverage ratio > 1
  cover = matrix(0, nrow = 1, ncol = m)
  for (i in 1:m) {
    if (wealth[nrow(wealth),i]<0) {
      cover[i] = sum(wealth[,i]>0)/n
    }
    else {
      cover[i] = (n+(wealth[nrow(wealth),i]/init))/n
    }
  }
  
  cover_ratio = mean(cover)
  
  return(cover)
}
```

Function 4:
input: 1. given return paths of equity, fixed income, inflation for the period of time
       2. asset allocation glidepath
       3. retirement fund starting balance and withdraw rate
       4. number of paths

output: coverage ratio distribution for given glidepath and SORR
```{r setup, include=FALSE}
SORR_to_coverage_dist = function(path_equity,path_FI,path_Inflation,glidepath_one,e,x,n,m) {
  withdraw = matrix(0,nrow = n, ncol = m)
  
  init = e*x/12
  withdraw[1,] = init
  
  for (i in 2:nrow(dt)) {
    for (j in 1:ncol(path_Inflation)) {
      withdraw[i,j] = withdraw[i-1,j] * (1+path_Inflation[i,j])
    }
  }
  
  wealth = matrix(0,nrow = n, ncol = m)
  wealth[1,] = e
  
  #calculate wealth position at specific time point after withdraw
  for (j in 1:m) {
    for (i in 2:n) {
      wealth[i,j] = (wealth[i-1,j] * ((1+path_equity[i,j])*glidepath[i,1] + (1+path_FI[i,j])*glidepath[i,2])) - withdraw[i,j]   
    }
  }
  
  
  # calculate coverage ratio for every path
  # if wealth at 30 year is still positive, coverage ratio > 1
  cover = matrix(0, nrow = 1, ncol = m)
  for (i in 1:m) {
    if (wealth[nrow(wealth),i]<0) {
      cover[i] = sum(wealth[,i]>0)/n
    }
    else {
      cover[i] = (n+(wealth[nrow(wealth),i]/init))/n
    }
  }
  
  #cover_ratio = mean(cover)
  
  return(cover)
}
```

Function 5: 
coverage distribution to Utility distribution 

```{r setup, include=FALSE}
coverage_dist_to_utility_dist = function(coverage_ratio, gamma, lambda,m) {
  utility = matrix(0,nrow = 1, ncol = m)
  for (i in 1:length(coverage_ratio)){
  if (coverage_ratio[1,i] >= 1){
    utility[1,i] = (coverage_ratio[1,i]^(1 - gamma) - 1) / (1 - gamma)
  }
  else{
    utility[1,i] = ((1^(1 - gamma) - 1) / (1 - gamma)) - lambda*(1 - coverage_ratio[1,i])
  }
  }
  return(utility)
}
```

Function 6:
mean Coverage ratio to utility

```{r setup, include=FALSE}
coverage_to_utility = function(coverage_ratio, gamma, lambda) {
  if (coverage_ratio >= 1){
    utility = (coverage_ratio^(1 - gamma) - 1) / (1 - gamma)
  }
  else{
    utility = ((1^(1 - gamma) - 1) / (1 - gamma)) - lambda*(1 - coverage_ratio)
  }
  
  return(utility)
}


```

General inputs:

```{r setup, include=FALSE}
num_path = 10000

SORR_risk = 0.5

initial_value_retirement = 1000000

percent_year_withdrawal = 0.04

gamma_input = 0.9999

lambda_input = 10

portion_equity = 0.6
portion_equity_start = 0.6 # 100-65
portion_equity_end = 0.3 # 100-95

coverage_or_utility = 'utility' # this should either equal to coverage or utility depending on what you would like to plot (will give you different graphs (everything will be automatic))
```


Run code

```{r setup, include=FALSE}
# simulate monte carlo paths for given SORR
gene_path <- Simulated_Path_Given_SORR(SORR_risk,
                                       num_path)
path_equity <- gene_path[[1]]
path_FI <- gene_path[[2]]
path_Inflation <- gene_path[[3]]

# generate glidepath
glidepath = gene_glidepath(mode = 2, start=portion_equity_start, end = portion_equity_end)

################## NOT NECESSARY ########################
#test on different allocation
#results = matrix(0,nrow=1,ncol=100)
#for (i in 1:100) {
#  a = i/100
#  results[i] = SORR_to_coverage(path_equity,path_FI,a,0.2,0.5,100000,0.04,342,100)
#}
#hist(results)
#plot(results[1,],type='l')
#########################################################
#plot coverage ratio histogram for given SORR and glidepath

results_hist = SORR_to_coverage_dist(path_equity = path_equity,
                                     path_FI = path_FI,
                                     path_Inflation = path_Inflation,
                                     glidepath = glidepath,
                                     e = initial_value_retirement,
                                     x = percent_year_withdrawal,
                                     n = 360,
                                     m = num_path)

results_utility = coverage_dist_to_utility_dist(results_hist,
                                           gamma = gamma_input,
                                           lambda = lambda_input,
                                           m = num_path)

if (coverage_or_utility == 'coverage'){
  median_hist = round(mean(results_hist),2)
  sd_hist = round(sd(results_hist),2)
  skew_hist = round(skewness(t(results_hist)),2)
  kurtosis_hist = round(kurtosis(t(results_hist)),2)
  hist(results_hist,
     breaks = 50,
     xlab = 'Coverage Ratio',
     freq = FALSE,
     main = paste(c('Mean = ', median_hist,
                    ' / SD = ', sd_hist,
                    ' / Skew = ', skew_hist,
                    ' / Kurtosis = ', kurtosis_hist),
                  collapse = ''))
} else{
  median_utility = round(mean(results_utility),2)
  sd_utility = round(sd(results_utility),2)
  skew_utility = round(skewness(t(results_utility)),2)
  kurtosis_utility = round(kurtosis(t(results_utility)),2)
  hist(results_utility,
     breaks = 50,
     xlab = 'Utility Coverage Ratio',
     freq = FALSE,
     main = paste(c('Mean = ', median_utility,
                    ' / SD = ', sd_utility,
                    ' / Skew = ', skew_utility,
                    ' / Kurtosis = ', kurtosis_utility),
                  collapse = ''))
}
```

CPPI strategg
input: return paths for equity, fixed income and inflation
output: utility or coverage graph for CPPI Strategy


```{r setup, include=FALSE}
library(FinancialMath)

CPPI_strategy = function(yearly_floor, time_years, interest_rate, path_equity,path_FI,
                         path_Inflation, e,x,n,m) {
  
  monthly_floor = (yearly_floor / 12)
  time_months = time_years * 12
  
  floor_df = matrix(0, nrow = time_months, ncol = 1)
  for (i in 1:time_months){
    floor_df[i,1] = NPV(cf0 = 0,
                        cf = as.vector(matrix(monthly_floor, nrow = i, ncol = 1)),
                        times = as.vector(matrix(seq(1,i,1), nrow= i, ncol = 1)),
                        i = interest_rate,
                        plot = FALSE)
  }
  rev_floor_df = (rev(floor_df))
  
  #find inverse return as a coefficient for equity weight
  inverse_return = matrix(0,nrow=1,ncol=m)
  for (i in 1:m) {
    inverse_return[i] = 1/min(path_equity[,i])
  }
  
  #find withdraw for each period
  withdraw = matrix(0,nrow = n, ncol = m)
  
  init = e*x/12
  withdraw[1,] = init
  
  for (i in 2:nrow(dt)) {
    for (j in 1:ncol(path_Inflation)) {
      withdraw[i,j] = withdraw[i-1,j] * (1+path_Inflation[i,j])
    }
  }
  
  #find wealth for each period
  wealth = matrix(0,nrow = n, ncol = m)
  C = matrix(0,nrow = n, ncol = m)
  dollar_eq = matrix(0,nrow = n, ncol = m)
  
  wealth[1,] = e
  C[1,] = e - rev_floor_df[1]
  
  #calculate wealth position at specific time point after withdraw
  for (j in 1:m) {
    dollar_eq[1,m] = inverse_return[j] * C[1,j]
    for (i in 2:n) {
      C[i,j] = wealth[i,j] - rev_floor_df[i]
      dollar_eq[i,j] = inverse_return[j] * C[i,j]
      wealth[i,j] = (wealth[i-1,j] * ((1+path_equity[i,j])*dollar_eq[i-1,j]/wealth[i-1,j] + (1+path_FI[i,j])*(1-dollar_eq[i-1,j]/wealth[i-1,j])))-withdraw[i,j]  
    }
  }
  
  cover = matrix(0, nrow = 1, ncol = m)
  for (i in 1:m) {
    if (wealth[nrow(wealth),i]<0) {
      cover[i] = sum(wealth[,i]>0)/n
    }
    else {
      cover[i] = (n+(wealth[nrow(wealth),i]/init))/n
    }
  }
  
  return(cover)
}


results_hist = CPPI_strategy(yearly_floor = 40000,
                             time_years = 30,
                             interest_rate = 0.1,
                             path_equity = path_equity,
                             path_FI = path_FI,
                             path_Inflation = path_Inflation,
                             e = 1000000,
                             x = 0.04,
                             n = 360,
                             m = 10000)

results_utility = coverage_dist_to_utility_dist(results_hist,
                                                gamma = gamma_input,
                                                lambda = lambda_input,
                                                m = num_path)

```



Show Income Effect depending on SORR

```{r setup, include=FALSE}

################################## Code for plotting - DO NOT RUN UNLESS NECESSARY #############################
# for output_num: columns are the 10 income levels, rows are the 10 SORR risk
# assumption: yearly withdrawal is 40,000
output_num = matrix(0,nrow = 10, ncol = 10)
for (i in 1:10){
  # generate the paths
  gene_path <- Simulated_Path_Given_SORR((i / 10),
                                       num_path)
  path_equity <- gene_path[[1]]
  path_FI <- gene_path[[2]]
  path_Inflation <- gene_path[[3]]
  for (j in 1:10){
    # generate the number
    output_num[i,j] = SORR_to_coverage_fix(path_equity = path_equity,
                                  path_FI = path_FI,
                                  path_Inflation = path_Inflation,
                                  a = portion_equity,
                                  start_point=0,
                                  end_point=0,
                                  e = j * 100000,
                                  n = 360,
                                  m = num_path)
  }
}

matplot(t(output_num),
        type = c("b"),
        pch=1,
        col = 1:10,
        xlab = 'Income (in $100k)',
        ylab = 'Coverage Ratio',
        main = 'With Annual Withdrawal of $40,000 p.a.')
```