---
title: "main_v5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Read in data

```{r setup, include=FALSE}
library(readxl)
filepath_ellie = "/Users/EllieLiu/Desktop/Monthly_returns.xlsx"
filepath_haocheng = '/Users/haochengtang/Desktop/MIT/Fall2018/Proseminar/data/clean\ data/Monthly_returns.xlsx'
#'/Users/hermannviktor/Python/PycharmProjects/15.451-Proseminar-IM/Excel/Monthly_returns.xlsx'

dt1 <- read_excel(filepath_haocheng)

```

Function 1:
input: 1. hitorical return data
       2. SORR 
       3. number of paths needed

output: return paths for equity and equity
        path_equity and path_FI are matrices with 342 rows (days of period) and n columns (number of path)
        
```{r setup, include=FALSE}
dt = dt1[, c(2,3,4)]
dt = dt[sample(1:nrow(dt), size = 360, replace = FALSE),]
colnames(dt) <- c("equity","FI","Inflation")

base = dt[order(dt$equity),]
n = nrow(base) 
equity_return = base$equity
FI_return = base$FI
Inflation = base$Inflation

size_bad_return = floor(0.3*n)
size_ok_return = n-size_bad_return
Simulated_Path_Given_SORR =
  function(sorr,m){
    index_bad = 1:floor(0.3*n) #108 periods
    index_ok = (floor(0.3*n)+1):n
    rand_index = matrix(0, nrow = n, ncol = m)
    path_equity = matrix(0, nrow = n, ncol = m)
    path_FI = matrix(0, nrow = n, ncol = m)
    path_Inflation = matrix(0, nrow = n, ncol = m)
    
    for(i in 1:m){
      left_bad = sample(index_bad, size = floor(sorr*size_bad_return))
      right_bad = setdiff(index_bad, left_bad)
      left_ok = sample(index_ok, size = (1/2*n - floor(sorr*size_bad_return)))
      right_ok = setdiff(index_ok, left_ok)
      left = c(left_bad, left_ok)
      right = c(right_bad, right_ok)
      
      #rand_index[1:(1/2*size_bad_return),i] = left_bad
      #rand_index[(1/2*n+1):(1/2*n+sorr*size_bad_return),i] = right_bad
      #rand_index[(1/2*size_bad_return+1):(1/2*n),i] = left_ok
      #rand_index[(1/2*n+1/2*size_bad_return+1):n,i] = right_ok
      
      rand_index[(1:(1/2*n)),i] <- sample(left, size = length(left))
      rand_index[((1/2*n+1):n),i] <- sample(right, size = length(right))
    }
    
    for (i in 1:m) {
      for (j in 1:n) {
        path_equity[j,i] = equity_return[rand_index[j,i]]
        path_FI[j,i] = FI_return[rand_index[j,i]]
        path_Inflation[j,i] = Inflation[rand_index[j,i]]
      }
    }
    return(list(path_equity,path_FI,path_Inflation))
  }

gene_path <- Simulated_Path_Given_SORR(0.5,100)
path_equity <- gene_path[[1]]
path_FI <- gene_path[[2]]
path_Inflation <- gene_path[[3]]
```

Function 2:
input: 1. given return paths of equity, fixed income, inflation for the period of time
       2. asset allocation glidepath
       3. retirement fund starting balance and withdraw rate
       4. number of paths

output: average coverage ratio for given glidepath and SORR

```{r setup, include=FALSE}
SORR_to_coverage = function(path_equity,path_FI,path_Inflation,a,start_point=0,end_point=0,e,x,n,m) {
  withdraw = matrix(0,nrow = n, ncol = m)
  #first glidepath, constant
  glidepath_one = matrix(0,nrow = n, ncol = 2)
  glidepath_one[,1] = a
  glidepath_one[,2] = 1-a
  
  #second glidepath
  glidepath_two = matrix(0,nrow = n, ncol = 2)
  rate = (end_point - start_point)/n  #changing rate
  glidepath_two[,1] = seq(from = start_point, to = end_point-rate, by = rate)
  glidepath_two[,2] = 1-glidepath_two[,1]
  
  init = e*x/12
  withdraw[1,] = init
  
  for (i in 2:nrow(dt)) {
    for (j in 1:ncol(path_Inflation)) {
      withdraw[i,j] = withdraw[i-1,j] * (1+path_Inflation[i,j])
    }
  }
  
  wealth = matrix(0,nrow = n, ncol = m)
  wealth[1,] = e
  
  #calculate wealth position at specific time point after withdraw
  for (j in 1:m) {
    for (i in 2:n) {
      wealth[i,j] = (wealth[i-1,j] * ((1+path_equity[i,j])*glidepath_one[i,1] + (1+path_FI[i,j])*glidepath_one[i,2])) - withdraw[i,j]   #change glide path to 2
    }
  }
  
  # calculate coverage ratio for every path
    # if wealth at 30 year is still positive, coverage ratio > 1
  cover = matrix(0, nrow = 1, ncol = m)
  for (i in 1:m) {
    if (wealth[nrow(wealth),i]<0) {
      cover[i] = sum(wealth[,i]>0)/n
    }
    else {
      cover[i] = (n+(wealth[nrow(wealth),i]/init))/n
    }
  }
  
  cover_ratio = mean(cover)
  
  return(cover_ratio)
}
```

Function 3:
input: 1. given return paths of equity, fixed income, inflation for the period of time
       2. asset allocation glidepath
       3. retirement fund starting balance and withdraw rate
       4. number of paths

output: coverage ratio distribution for given glidepath and SORR
```{r setup, include=FALSE}
SORR_to_coverage_dist = function(path_equity,path_FI,path_Inflation,a,start_point=0,end_point=0,e,x,n,m) {
  withdraw = matrix(0,nrow = n, ncol = m)
  
  #first glidepath, constant
  glidepath_one = matrix(0,nrow = n, ncol = 2)
  glidepath_one[,1] = a
  glidepath_one[,2] = 1-a
  
  #second glidepath
  glidepath_two = matrix(0,nrow = n, ncol = 2)
  rate = (end_point - start_point)/n  #changing rate
  glidepath_two[,1] = seq(from = start_point, to = end_point-rate, by = rate)
  glidepath_two[,2] = 1-glidepath_two[,1]
  
  init = e*x/12
  withdraw[1,] = init
  
  for (i in 2:nrow(dt)) {
    for (j in 1:ncol(path_Inflation)) {
      withdraw[i,j] = withdraw[i-1,j] * (1+path_Inflation[i,j])
    }
  }
  
  wealth = matrix(0,nrow = n, ncol = m)
  wealth[1,] = e
  
  #calculate wealth position at specific time point after withdraw
  for (j in 1:m) {
    for (i in 2:n) {
      wealth[i,j] = (wealth[i-1,j] * ((1+path_equity[i,j])*glidepath_one[i,1] + (1+path_FI[i,j])*glidepath_one[i,2])) - withdraw[i,j]   #change glide path to 2
    }
  }
  
  
  # calculate coverage ratio for every path
  # if wealth at 30 year is still positive, coverage ratio > 1
  cover = matrix(0, nrow = 1, ncol = m)
  for (i in 1:m) {
    if (wealth[nrow(wealth),i]<0) {
      cover[i] = sum(wealth[,i]>0)/n
    }
    else {
      cover[i] = (n+(wealth[nrow(wealth),i]/init))/n
    }
  }
  
  #cover_ratio = mean(cover)
  
  return(cover)
}
```

Function: Utility coverage

```{r setup, include=FALSE}
coverage_to_utility_dist = function(coverage_ratio, gamma, lambda,m) {
  utility = matrix(0,nrow = 1, ncol = m)
  for (i in 1:length(coverage_ratio)){
  if (coverage_ratio[1,i] >= 1){
    utility[1,i] = (coverage_ratio[1,i]^(1 - gamma) - 1) / (1 - gamma)
  }
  else{
    utility[1,i] = ((1^(1 - gamma) - 1) / (1 - gamma)) - lambda*(1 - coverage_ratio[1,i])
  }
  }
  return(utility)
}
```

Run code

```{r setup, include=FALSE}
gene_path <- Simulated_Path_Given_SORR(0.3,100000)
path_equity <- gene_path[[1]]
path_FI <- gene_path[[2]]
path_Inflation <- gene_path[[3]]

################## NOT NECESSARY ########################
#test on different allocation
#results = matrix(0,nrow=1,ncol=100)
#for (i in 1:100) {
#  a = i/100
#  results[i] = SORR_to_coverage(path_equity,path_FI,a,0.2,0.5,100000,0.04,342,100)
#}
#hist(results)
#plot(results[1,],type='l')
#########################################################
#plot coverage ratio histogram for given SORR and glidepath
results_hist = SORR_to_coverage_dist(path_equity = path_equity,path_FI = path_FI, path_Inflation = path_Inflation, a=0.6,e=100000,x=0.04,n=360,m=100000)
results_utility = coverage_to_utility_dist(results_hist,gamma = 0.9999, lambda = 10, m = 100000)
median_utility = round(mean(results_utility),2)
sd_utility = round(sd(results_utility),2)
#hist(results_hist, breaks = 12, xlab = 'Coverage Ratio', main = 'SORR=0.7 / Equity proportion = 0.6')
hist(results_utility, breaks = 12, xlab = 'Utility Coverage Ratio', main = paste(c('Mean = ',  median_utility, ' / Standard deviation = ', sd_utility), collapse = ''))
```